\documentclass[aspectratio=169,11pt]{beamer}

\usetheme{default}
\setbeamertemplate{navigation symbols}{}
\setbeamertemplate{headline}{}
\setbeamersize{text margin left=0.85cm,text margin right=0.85cm}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[scaled=0.95]{helvet}
\renewcommand{\familydefault}{\sfdefault}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{array}
\usepackage{hyperref}
\usepackage{tikz}

\definecolor{datamakOrange}{HTML}{EA7B17}
\definecolor{datamakCharcoal}{HTML}{222222}
\definecolor{datamakGray}{HTML}{5A5A5A}
\definecolor{datamakLight}{HTML}{F5F7FA}
\definecolor{datamakBorder}{HTML}{D9E2EC}
\definecolor{datamakBlue}{HTML}{0B3D91}
\definecolor{datamakGreen}{HTML}{2B6E46}
\definecolor{datamakRed}{HTML}{B42318}

\setbeamercolor{normal text}{fg=datamakCharcoal,bg=white}
\setbeamercolor{structure}{fg=datamakOrange}
\setbeamercolor{block title}{fg=datamakCharcoal,bg=datamakBorder}
\setbeamercolor{block body}{fg=datamakCharcoal,bg=datamakLight}
\setbeamercolor{itemize item}{fg=datamakOrange}
\setbeamercolor{itemize subitem}{fg=datamakOrange}
\setbeamercolor{enumerate item}{fg=datamakOrange}
\setbeamerfont{title}{size=\LARGE,series=\bfseries}
\setbeamerfont{subtitle}{size=\large,series=\bfseries}
\setbeamerfont{frametitle}{size=\large,series=\bfseries}
\setbeamerfont{block title}{size=\normalsize,series=\bfseries}
\setbeamerfont{footline}{size=\scriptsize}

\setbeamertemplate{blocks}[default]
\setbeamertemplate{itemize item}{\color{datamakOrange}$\blacktriangleright$}
\setbeamertemplate{itemize subitem}{\color{datamakOrange}\scalebox{0.8}{$\blacktriangleright$}}

\makeatletter
\setbeamertemplate{frametitle}{%
  \nointerlineskip
  \begin{tikzpicture}[remember picture,overlay]
    \fill[datamakOrange] (current page.north west) rectangle ([yshift=-1.02cm]current page.north east);
    \node[
      anchor=north,
      text=white,
      font=\usebeamerfont{frametitle},
      yshift=-0.18cm
    ] at (current page.north) {\strut\insertframetitle};
  \end{tikzpicture}%
  \vspace*{0.88cm}
  \vspace{0.18cm}
}
\makeatother

\setbeamertemplate{footline}{%
  \begin{tikzpicture}[remember picture,overlay]
    \node[
      anchor=south east,
      text=datamakGray,
      font=\usebeamerfont{footline}
    ] at ([xshift=-0.22cm,yshift=0.10cm]current page.south east) {\insertframenumber};
  \end{tikzpicture}%
}

\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\lite}{Datamak Lite}
\newcommand{\relation}[1]{\textcolor{datamakBlue}{\texttt{#1}}}

\title{\lite}
\subtitle{A lightweight campaign-visibility layer for scientific simulations}
\author{}
\date{\today}

\begin{document}

\begin{frame}[plain]
  \vspace*{0.55cm}
  \begin{center}
    \begin{tikzpicture}
      \node[
        fill=datamakOrange,
        draw=datamakCharcoal!45,
        line width=0.6pt,
        text=white,
        minimum width=0.96\textwidth,
        minimum height=2.35cm,
        text width=0.82\textwidth,
        align=center,
        inner xsep=0.6cm,
        inner ysep=0.45cm
      ] {\bfseries\fontsize{27}{32}\selectfont \inserttitle};
    \end{tikzpicture}
  \end{center}
  \vspace{0.8cm}
  \begin{center}
    {\usebeamerfont{subtitle}\insertsubtitle\par}
    \vspace{0.4cm}
    {\large Generic campaign visibility and provenance model\par}
  \end{center}
\end{frame}

\begin{frame}{Why Lite Exists}
  \begin{columns}[T]
    \column{0.48\textwidth}
    \begin{block}{The problem}
      \small
      A campaign quickly becomes a graph, not a list:
      \begin{itemize}
        \item simulations produce reusable datasets;
        \item pools, restarts, and analyses reuse or compare against them;
        \item figures mix old and new results;
        \item decisions evolve as bugs and assumptions are found.
      \end{itemize}
    \end{block}

    \column{0.48\textwidth}
    \begin{block}{The goal}
      \small
      Lite should answer practical questions:
      \begin{itemize}
        \item What produced this figure?
        \item Which input dataset did this pool use?
        \item Which result is the current reference?
        \item What is running, pending, partial, or suspect?
        \item Why did we make this choice?
      \end{itemize}
    \end{block}
  \end{columns}
\end{frame}

\begin{frame}{Core Data Model}
  \begin{block}{Five generic objects}
    \small
    \begin{tabular}{>{\bfseries}p{0.18\linewidth}p{0.72\linewidth}}
      \toprule
      Entity & Simulation, pool, history file, restart file, analysis, figure, slide, paper result, campaign. \\
      Relation & Directed link: \relation{uses\_history}, \relation{restart\_from}, \relation{produces}, \relation{plots}, \relation{compares\_to}, \relation{supersedes}. \\
      Artifact & Concrete file or directory path attached to an entity. \\
      Metric & Compact scalar result or parameter, such as \code{D}, \code{PF}, \code{dt}, \code{time\_end}, or \code{effective\_stride}. \\
      Note & Markdown comment, warning, todo, or scientific decision. \\
      \bottomrule
    \end{tabular}
  \end{block}

  \begin{block}{Status separation}
    \small
    Operational status says whether the work ran.  Scientific status says
    whether the result should be trusted, used, superseded, or treated as
    suspect.
  \end{block}
\end{frame}

\begin{frame}{The Sidecar Contract}
  \begin{columns}[T]
    \column{0.45\textwidth}
    \begin{block}{Local provenance receipt}
      \small
      Every important object should leave:
      \begin{center}
        \code{run\_or\_pool/datamak\_lite.json}
      \end{center}
      The sidecar records paths, relations, compact metrics, and notes.
    \end{block}

    \column{0.51\textwidth}
    \begin{block}{Why this matters}
      \small
      \begin{itemize}
        \item It is visible next to the simulation.
        \item It is editable by a human or an agent.
        \item It can rebuild the SQLite campaign view.
        \item It avoids copying large history, restart, or NetCDF files.
      \end{itemize}
    \end{block}
  \end{columns}

  \begin{block}{Minimal workflow}
    \small
    Prepare object $\rightarrow$ write sidecar $\rightarrow$ import SQLite
    registry $\rightarrow$ inspect report or GUI $\rightarrow$ add decisions.
  \end{block}
\end{frame}

\begin{frame}{Human And Agent Workflow}
  \begin{block}{During normal campaign work}
    \small
    When a simulation pool, restart chain, analysis, or figure is prepared, the
    agent should also create or update the Lite metadata.  This is not a
    separate bookkeeping session; it is part of the workflow.
  \end{block}

  \begin{block}{Example command for a generic pool}
    \scriptsize
    \code{python3 -m datamak\_lite.cli create-pool-packet POOL\_ROOT
    --campaign-uid CAMPAIGN\_UID
    --uses-dataset-uid DATASET\_UID
    --dataset-path /path/to/input\_dataset
    --import-db /path/to/campaign.sqlite
    --report}
  \end{block}

  \begin{block}{Remote rule}
    \small
    For Perlmutter, copy or sync only the small sidecar metadata.  Do not copy
    large simulation data merely to update Lite.
  \end{block}
\end{frame}

\begin{frame}{What The GUI Should Show}
  \begin{columns}[T]
    \column{0.48\textwidth}
    \begin{block}{First useful GUI}
      \small
      \begin{itemize}
        \item campaign object list;
        \item selected entity details;
        \item ``Depends On'' and ``Used By'' panels;
        \item artifacts, metrics, notes, and warnings;
        \item compact status chips.
      \end{itemize}
    \end{block}

    \column{0.48\textwidth}
    \begin{block}{Design constraint}
      \small
      The GUI is a frontend only.  The same SQLite database and core API must
      be usable from CLI scripts and agents.
    \end{block}

    \begin{block}{Visual identity}
      \small
      Reuse Datamak's compact interface: Helvetica/Arial, restrained
      orange/gray accents, white panels, dense tables, and calm status colors.
    \end{block}
  \end{columns}
\end{frame}

\begin{frame}{Relationship With Datamak}
  \begin{block}{Datamak}
    \small
    Datamak is a workflow system: database-backed input generation, batch
    creation, remote launch and monitoring, result sync, and surrogate-model
    workflows for simulation and analysis studies.
  \end{block}

  \begin{block}{Lite}
    \small
    Lite is a campaign visibility system: provenance, relation graph, figure
    links, scientific notes, decisions, and operational/scientific status across
    multiple codes.
  \end{block}

  \begin{block}{Complementarity}
    \small
    Datamak-style pools can own execution state.  Lite records how those pools
    connect to histories, restarts, analyses, figures, and decisions.
  \end{block}
\end{frame}

\begin{frame}{Relationship With ORNL Campaign Tools}
  \begin{block}{Campaign management direction}
    \small
    ORNL's Workflow Systems group frames campaign management as coordinated
    planning, collection, processing, storage, and analysis of scientific data.
    Lite is aligned with that direction, but starts with the local campaign
    graph and day-to-day scientific decisions.
  \end{block}

  \begin{block}{Complementary tools}
    \small
    \begin{tabular}{>{\bfseries}p{0.20\linewidth}p{0.69\linewidth}}
      \toprule
      DataFed & Open-source, domain-agnostic scientific data federation for managing, sharing, and discovering working data across facilities. \\
      ADIOS2/HPC Campaign & Toolkit for `.ACA' campaign archives with metadata for ADIOS2/HDF5 outputs, images, and text files. \\
      Lite & Local, lightweight campaign brain: graph, sidecars, notes, decisions, figures, and code-agnostic status. \\
      \bottomrule
    \end{tabular}
  \end{block}
\end{frame}

\begin{frame}{CMSD, DataFed, And ADIOS2/HPC Campaign}
  \begin{block}{Positioning}
    \small
    CMSD is not a single obvious competitor to Lite.  It is better understood
    as a broader ORNL/DOE concept and tool family around campaign-scale
    scientific data management.  ORNL lists CMSD alongside tools such as ADIOS,
    MGARD, SimGrid, and EFFIS in its Workflow Systems work.
  \end{block}

  \begin{columns}[T]
    \column{0.48\textwidth}
    \begin{block}{ADIOS2 / HPC Campaign}
      \scriptsize
      Creates small `.ACA' campaign archives pointing to large ADIOS2/HDF5,
      image, and text datasets, with metadata and file locations.  This is
      close to Lite's sidecar idea, but is more data-file and remote-access
      oriented.  The ADIOS2 documentation notes that campaign management is
      still fairly new and expected to evolve.
    \end{block}

    \column{0.48\textwidth}
    \begin{block}{DataFed}
      \scriptsize
      Open-source ORNL scientific data federation.  It is more ambitious than
      Lite: records, metadata, relationships, services, web portal, CLI,
      Globus-based data movement, and cross-facility access.  Lite should be
      able to store DataFed record or collection IDs later.
    \end{block}
  \end{columns}
\end{frame}

\begin{frame}{Interoperability Roadmap}
  \begin{block}{Do not pivot yet}
    \small
    The first version should not depend on DataFed, ADIOS2 campaign archives,
    Globus, or a facility service.  The immediate value is the local
    agent-maintained campaign graph.
  \end{block}

  \begin{block}{Keep external hooks}
    \small
    Later, Lite can store external references as artifacts or metadata:
    \begin{itemize}
      \item DataFed record or collection id;
      \item Globus endpoint and path;
      \item ADIOS2/HPC Campaign `.ACA' path;
      \item DOI or publication archive id;
      \item external workflow-system run id.
    \end{itemize}
  \end{block}
\end{frame}

\begin{frame}{Implementation Roadmap}
  \begin{block}{Current stage}
    \small
    \begin{itemize}
      \item generic SQLite schema;
      \item packet import and remote sidecar sync;
      \item registry, folder-inventory, and figure-audit adapters;
      \item first read-only GUI.
    \end{itemize}
  \end{block}

  \begin{block}{Next useful steps}
    \small
    \begin{enumerate}
      \item Import figure JSON metadata so figures become first-class graph nodes.
      \item Improve the GUI around figure-driven questions.
      \item Add code-specific sidecar creators only after one real object proves the need.
      \item Add external archive/catalog fields when the local workflow is stable.
    \end{enumerate}
  \end{block}
\end{frame}

\begin{frame}{References}
  \small
  \begin{itemize}
    \item ORNL Workflow Systems group:
      \url{https://www.ornl.gov/group/workflow-systems}
    \item DataFed introduction:
      \url{https://ornl.github.io/DataFed/system/introduction.html}
    \item DataFed system overview:
      \url{https://ornl.github.io/DataFed/system/overview.html}
    \item ADIOS2/HPC Campaign Management:
      \url{https://adios2.readthedocs.io/en/latest/advanced/campaign_management.html}
    \item HPC Campaign documentation:
      \url{https://hpc-campaign.readthedocs.io}
    \item DataFed GitHub:
      \url{https://github.com/ORNL/DataFed}
  \end{itemize}
\end{frame}

\end{document}
